Robust calibration of numerical models based on relative regret

### 2019-11-19 article-modif-A&E.tex
estimation on swe

Given that $J>0$, $\Gamma_{\alpha}(\mathbf{k})$ is increasing with respect to $\alpha$ for any $\mathbf{k}\in\Kspace$. We can then focus on the smallest value of $\alpha$ such that $\Gamma_\alpha$ reaches a certain level of confidence $p\in[0,1]$. This leads to the definition of $\checkap$
\begin{align}
  \checkap &= \inf\left\{ \alpha\geq 1 \mid \exists \checkkp \in \Kspace,\, \Gamma_{\alpha}(\checkkp) \geq p \right\} \nonumber \\
   &= \inf \left\{ \alpha \geq 1 \mid \max_{\mathbf{k}\in\Kspace} \Gamma_{\alpha}(\mathbf{k}) \geq p \right\}   \label{eq:def_alpha_check}
\end{align}
that is the smallest $\alpha$, such that there exists a particular $\checkkp \in \Kspace$ for which $J(\checkkp,\mathbf{U}) \leq \checkap J^*(\mathbf{U})$ with probability $p$. This formulation can be linked to the quantiles and Value-at-Risk of the random variable $\max_{\mathbf{k}} \{ \frac{J(\mathbf{k},\mathbf{U})}{J^*(\mathbf{U})}\}$, that is a measure of risk usually applied in the financial sector (see~\cite{rockafellar_deviation_2002}). For different levels $p$, and thus different $\alpha_p$, Figure~\ref{fig:illu_alpha_p} shows examples of $\Gamma_{\alpha_p}$. By definition, the associated $\mathbf{k}_p$ is then the first value for which $\Gamma_{\alpha_p}$ reaches $p$. We can see that changing the level $p$ shifts the maximiser of $\Gamma_{\alpha_p}$.

@@

\begin{table*}[t]
  \centering
\label{tab:RO_recap}
\begin{tabular}{lrr}
  \toprule
  Definition & Related quantities & Interpretation \\ \midrule
   $\argmin_{\mathbf{k}\in\Kspace} \Ex_{\mathbf{U}}\left[J(\mathbf{k},\mathbf{U})\right]$& $\kmean$ & Long run performances\\
   $\argmin_{\mathbf{k}\in\Kspace} \Var_{\mathbf{U}}\left[J(\mathbf{k},\mathbf{U})\right]$& $\kvar$  & Steady performances\\
   $\argmax_{\mathbf{k}\in\Kspace} p_{\mathbf{K}^*}(\mathbf{k})$ & $\kmpe$ & Most probable minimiser\\
    $\inf\left\{ \alpha \mid \exists \checkkp \in \Kspace,\, \Gamma_{\alpha}(\checkkp) \geq p \right\}$ & $(p, \checkkp,\checkap)$ & Acceptable values with fixed probability $p$ \\
    $p_{\mathrm{maxratio}} = \argmax p/\checkap$ & $(p_{\mathrm{maxratio}}, \checkk_{\mathrm{maxratio}}, \checka_{\mathrm{maxratio}})$ & Maximal ratio of $p$ and $\checkap$ \\
  \bottomrule
\end{tabular}
  \caption{Robust estimators, based on a cost function $J$}

\end{table*}

@@

For the $J_{\mathrm{BH}}$ function, the curves are plotted in Figure~\ref{fig:profiles_branin}. Based on its definition, the profile of $\hat{\checkk}_1$ is always within the shaded region, corresponding to $[J^*(\mathbf{u}), \hat{\checka}_1 J^*(\mathbf{u})]$. The profile of $\hatkmean$ in contrast, exceeds $\hat{\checka}_1 J^*(\mathbf{u})$ for $\mathbf{u}$ close to $5$, while the profile of $\hatkvar$ does for $\mathbf{u}$ close to $0$.

@@

In this case, $\hatkmpe$, $\hatkmean$ and $\hat{\checkk}_{\mathrm{maxratio}}$ present a similar behaviour. They perform very well for $\mathbf{u}>2$, especially for $\hatkmpe$ which is very close to the minimal value; for $\mathbf{u}<2$, it produces high values of the function. 
The performances of $\hat{\checkk}_1$ are closer to the performances of $\hatkvar$ for this function, but it performs worse than $\hatkmean$ and $\hatkvar$ for $\mathbf{u}>2$, even though its range is designed to stay within the interval $[J^*(\mathbf{u}); \hat{\checka}_1 J^*(\mathbf{u})]$.

@@

\begin{align}
  \label{eq:SWE_equations}
  \left\{
    \begin{array}{rl}
      \partial_t h + \partial_x q &= 0 \\
      \partial_t q + \partial_x \left(\frac{q^2}{h} + \frac{g}{2}h^2\right) &= - gh\partial_x z - \mathbf{k}q |q|h^{-7/3} \\
      h(0, t) &= 20.0 + 3\cdot \sin \left(\frac{2\pi t}{2}\right) + 1.5\cdot \mathbf{u} \cdot \sin\left(\frac{2 \pi t}{\omega_0}\right)\\
      \partial_x q(0, t) &= 0
      \end{array}
  \right.
\end{align}
These equations are integrated using a finite-volume scheme on a discretized domain $[0, L]$, up to a time $T$. The output of the computer code is the sea surface height $h$, on the center of all the volumes and at all the time-steps, that will be denoted $\mathcal{M}(\mathbf{k},\mathbf{u};\, \omega_0=1.0)$.
In this setting, the variable input is uniformly distributed on $\mathcal{U}$.
To generate the observations, we choose $\mathbf{u}^{\mathrm{true}}=2/3$, and define $\mathcal{M}^o$ based on the computer model $\mathcal{M}$, such that $\mathcal{M}^o(\mathbf{u}^{\mathrm{true}}) = \mathcal{M}(\mathbf{k}^{\mathrm{true}}, \mathbf{u}^{\mathrm{true}}; \, \omega_0 = 0.999)$. $\omega_0$  
represents here the uncontrollable error between the observations and the numerical model and will now be omitted systematically in the notation. The true value of the bottom friction $\mathbf{k}^{\mathrm{true}} = (k_1^{\mathrm{true}}, k_2^{\mathrm{true}},\dots, k_{N_\mathrm{vol}}^{\mathrm{true}})$ is not constant over the whole domain, and is defined as
\begin{equation*}
{k}^{\mathrm{true}}_i = 0.2\cdot \left(1 + \sin\left(\frac{2 \pi x_i}{L}\right)\right)
\end{equation*}
where $x_i$ is the center of the $i$-th volume. The two sources of representativity errors are $\omega_0$, and the fact that we are looking to estimate $\mathbf{k}$ in a 1-dimensional space $\Kspace$. 
\subsection{Estimation}
Given that the numerical model is expensive to evaluate, it has first been evaluated on a relatively small regular grid. A metamodel based on Gaussian process is constructed and the PEI criterion \citep{ginsbourger_bayesian_2014,bossek_learning_2015} allow us to add point to evaluate and to better capture the locus of the conditional minimisers $\{(\mathbf{k}^*(\mathbf{u}), \mathbf{u}) \mid \mathbf{u} \in \Uspace \}$. A bigger regular grid is evaluated by the metamodel once the design space has been sufficiently explored. The different steps of the estimation are illustrated Figure~\ref{fig:estimation_swe}. We can see that the estimation for this problem seems less problematic than the analytical examples shown above, as all derived quantites are unimodal.
The different estimates however take a wide range of values, as seen on Table~\ref{tab:recap_estimates_metaSWE}.: there is a unique global minimiser on the whole space $\Kspace \times \Uspace$, noted $\kest_{\mathrm{global}}$. As discussed earlier, the estimation of the MPE is dependent on the method used, and then its estimation using KDE in the middle plot of Figure~\ref{fig:estimation_swe} is here only for illustrative purposes.

\begin{figure}[!h]
\centering
\includegraphics[width=.7\textwidth]{Figures/RobustSWEmis_plots.pdf}
\caption{Procedure of robust calibration for the shallow water problem. Top: contours of $J$, conditional minimisers and $\{(\mathbf{k},\mathbf{u}) | J(\mathbf{k},\mathbf{u}) = \hat{\alpha}_1 J^*(\mathbf{u})\}$. Middle: Conditional moments and histogram and KDE of the conditional minimisers. Bottom: $\hat{\Gamma}_{\alpha_p}$ for different levels $p$}
\label{fig:estimation_swe}
\end{figure}
\begin{table*}[!h]
\centering
\caption{Calibrated values of $\mathbf{k}$ according to different criteria}
\begin{tabular}{lr} \toprule
Estimator & Value \\ \midrule
$\hatkmpe$ & 0.249 \\
$\kest_{\mathrm{global}}$ & 0.290 \\
$\hat{\mathbf{k}}_{p},~p=1$ & 0.423 \\ 
$\hat{\mathbf{k}}_p,~p=0.90$ & 0.458 \\ 
$\hatkmean$ & 0.501 \\ 
$\hat{\mathbf{k}}_p,~p=0.80$ & 0.505 \\ 
$\hat{\mathbf{k}}_p,~p=0.70$ & 0.560 \\ 
$\hatkvar$ & 0.990\\ \bottomrule
\end{tabular}
\label{tab:recap_estimates_metaSWE}
\end{table*}

\begin{figure}[!h]
  \centering
  \includegraphics[width=.8\textwidth]{Figures/profile_swe}
  \caption{Profiles for the calibration problem}
\label{fig:profiles_swe}
\end{figure}

\subsection{Assessing the quality of the forecast of the calibrated model}
The calibration of the model has been performed using by integrating the model on a time-period $[0, T]$, called assimilation window. 
We now want to compare the quality of the different forecasts, from different calibrated bottom friction.
Those forecasts result from the integration of the numerical model between the time $T$ and a time $T_{\mathrm{pred}}$.

>>>>

Given that $J>0$, $\Gamma_{\alpha}(\mathbf{k})$ is increasing with respect to $\alpha$ for any $\mathbf{k}\in\Kspace$. We can then focus on the smallest value of $\alpha$ such that $\Gamma_\alpha$ reaches a certain level of confidence $p\in[0,1]$. This leads to the definition of $\checkap$
\begin{align}
  \checkap &= \inf\left\{ \alpha\geq 1 \mid \exists \checkkp \in \Kspace,\, \Gamma_{\alpha}(\checkkp) \geq p \right\} \nonumber \\
   &= \inf \left\{ \alpha \geq 1 \mid \max_{\mathbf{k}\in\Kspace} \Gamma_{\alpha}(\mathbf{k}) \geq p \right\}   \label{eq:def_alpha_check}
\end{align}
that is the smallest $\alpha$, such that there exists a particular $\checkkp \in \Kspace$ for which $J(\checkkp,\mathbf{U}) \leq \checkap J^*(\mathbf{U})$ with probability $p$. This formulation can be linked to the quantiles and Value-at-Risk of the random variable $\max_{\mathbf{k}} \{ \frac{J(\mathbf{k},\mathbf{U})}{J^*(\mathbf{U})}\}$, that is a measure of risk usually applied in the financial sector (see~\cite{rockafellar_deviation_2002}). For different levels $p$, and thus different $\alpha_p$, Figure~\ref{fig:illu_alpha_p} shows examples of $\Gamma_{\alpha_p}$. By definition, the associated $\mathbf{k}_p$ is then the first value for which $\Gamma_{\alpha_p}$ reaches $p$. We can see that changing the level $p$ shifts the maximiser of $\Gamma_{\alpha_p}$, and that for small $\alpha$, $\Gamma_{\alpha}(\mathbf{k}_1)$ is also very small. This indicates that $\mathbf{k}_1$ is located quite far from the conditional minimisers, and thus represents a compromise.

@@

\begin{table}[t]
  \centering
\begin{tabular}{lrr}
  \toprule
  Definition & Related quantities & Interpretation \\ \midrule
   $\argmin_{\mathbf{k}\in\Kspace} \Ex_{\mathbf{U}}\left[J(\mathbf{k},\mathbf{U})\right]$& $\kmean$ & Long run performances\\
   $\argmin_{\mathbf{k}\in\Kspace} \Var_{\mathbf{U}}\left[J(\mathbf{k},\mathbf{U})\right]$& $\kvar$  & Steady performances\\
   $\argmax_{\mathbf{k}\in\Kspace} p_{\mathbf{K}^*}(\mathbf{k})$ & $\kmpe$ & Most probable minimiser\\
    $\inf\left\{ \alpha \mid \exists \checkkp \in \Kspace,\, \Gamma_{\alpha}(\checkkp) \geq p \right\}$ & $(p, \checkkp,\checkap)$ & Acceptable values with fixed probability $p$ \\
    $p_{\mathrm{maxratio}} = \argmax p/\checkap$ & $(p_{\mathrm{maxratio}}, \checkk_{\mathrm{maxratio}}, \checka_{\mathrm{maxratio}})$ & Maximal ratio of $p$ and $\checkap$ \\
  \bottomrule
\end{tabular}
\label{tab:RO_recap}
  \caption{Robust estimators, based on a cost function $J$}
\end{table}

@@

For the $J_{\mathrm{BH}}$ function, the curves are plotted in Figure~\ref{fig:profiles_branin}. Based on its definition, the profile of $\hat{\checkk}_1$ is always within the shaded region, corresponding to $[J^*(\mathbf{u}), \hat{\checka}_1 J^*(\mathbf{u})]$. The profile of $\hatkmean$ in contrast, exceeds $\hat{\checka}_1 J^*(\mathbf{u})$ for $\mathbf{u}$ close to $5$, while the profile of $\hatkvar$ does for $\mathbf{u}$ close to $0$.
As the different $\kest$ are very close for every criteria introduced in this paper, the profiles are almost undistinguishable.

@@

In this case, $\hatkmpe$, $\hatkmean$ and $\hat{\checkk}_{\mathrm{maxratio}}$ present a similar behaviour. They perform very well for $\mathbf{u}>2$, especially for $\hatkmpe$ which is very close to the minimal value; However for $\mathbf{u}<2$, they produce high values of the function.

The performances of $\hat{\checkk}_1$ are closer to the performances of $\hatkvar$ for this function, but it performs worse than $\hatkmean$ and $\hatkvar$ for $\mathbf{u}>2$, even though its range is designed to stay within the interval $[J^*(\mathbf{u}); \hat{\checka}_1 J^*(\mathbf{u})]$.

@@

\begin{align}
  \label{eq:SWE_equations}
  \left\{
    \begin{array}{rl}
      \partial_t h + \partial_x q &= 0 \\
      \partial_t q + \partial_x \left(\frac{q^2}{h} + \frac{g}{2}h^2\right) &= - gh\partial_x z - \mathbf{k}q |q|h^{-7/3} \\
      h(0, t) &= 20.0 + 3\cdot \sin \left(\frac{2\pi t}{2}\right) + 1.5\cdot \mathbf{u} \cdot \sin\left(\frac{2 \pi t}{\omega_0}\right)\\
      \partial_x q(0, t) &= 0
      \end{array}
  \right.
\end{align}
These equations are integrated using a finite-volume scheme on a discretized domain $[0, L]$, up to a time $T$. The output of the computer code is the sea surface height $h$, on the center of all the volumes and at all the time-steps, that will be denoted $\mathcal{M}(\mathbf{k},\mathbf{u};\, \omega_0=1.0)$.
In this setting, the variable input is uniformly distributed on $\mathcal{U}$.
To generate the observations, we choose $\mathbf{u}^{\mathrm{true}}=2/3$, and define $\mathcal{M}^o$ based on the computer model $\mathcal{M}$, such that $\mathcal{M}^o(\mathbf{u}^{\mathrm{true}}) = \mathcal{M}(\mathbf{k}^{\mathrm{true}}, \mathbf{u}^{\mathrm{true}}; \, \omega_0 = 0.999)$. $\omega_0$  
represents here the uncontrollable error between the observations and the numerical model and will now be omitted systematically in the notation. The true value of the bottom friction $\mathbf{k}^{\mathrm{true}} = (k_1^{\mathrm{true}}, k_2^{\mathrm{true}},\dots, k_{N_\mathrm{vol}}^{\mathrm{true}})$ is not constant over the whole domain, and is defined as
\begin{equation*}
{k}^{\mathrm{true}}_i = 0.2\cdot \left(1 + \sin\left(\frac{2 \pi x_i}{L}\right)\right)
\end{equation*}
where $x_i$ is the center of the $i$-th volume. The two sources of representativity errors are $\omega_0$, and the fact that we are looking to estimate $\mathbf{k}$ in a 1-dimensional space $\Kspace$. 
\subsection{Estimation}
Given that the numerical model is expensive to evaluate, it has first been evaluated on a relatively small regular grid. A metamodel based on Gaussian process is constructed and the PEI criterion \citep{ginsbourger_bayesian_2014,bossek_learning_2015} allow us to add point to evaluate and to better capture the locus of the conditional minimisers $\{(\mathbf{k}^*(\mathbf{u}), \mathbf{u}) \mid \mathbf{u} \in \Uspace \}$. A bigger regular grid is evaluated by the metamodel once the design space has been sufficiently explored. The different steps of the estimation are illustrated Figure~\ref{fig:estimation_swe}. We can see that the estimation for this problem seems less problematic than the analytical examples shown above, as all derived quantites are unimodal.
The different estimates however take a wide range of values, as seen on Table~\ref{tab:recap_estimates_metaSWE}. As a basis for comparison, the global minimiser of $J$ over $\Kspace \times \Uspace$ has been computed, and $\kest_{\mathrm{global}}$ is then obtained by discarding the $\mathbf{u}$ value. An interesting comparison arises when comparing with $\kest_{\mathrm{global}}$, where $\kest_1$ outperforms the global minimiser for $\mathbf{u} < 0.625$.

As discussed earlier, the estimation of the MPE is dependent on the method used, and then its estimation using KDE in the middle plot of Figure~\ref{fig:estimation_swe} is here only for illustrative purposes.

\begin{figure}[!h]
\centering
\includegraphics[width=.7\textwidth]{Figures/RobustSWEmis_plots.pdf}
\caption{Procedure of robust calibration for the shallow water problem. Top: contours of $J$, conditional minimisers and $\{(\mathbf{k},\mathbf{u}) | J(\mathbf{k},\mathbf{u}) = \hat{\alpha}_1 J^*(\mathbf{u})\}$. Middle: Conditional moments and histogram and KDE of the conditional minimisers. Bottom: $\hat{\Gamma}_{\alpha_p}$ for different levels $p$}
\label{fig:estimation_swe}
\end{figure}
\begin{table*}[!h]
\centering
\caption{Calibrated values of $\mathbf{k}$ according to different criteria}
\begin{tabular}{lr} \toprule
Estimator & Value \\ \midrule
$\hatkmpe$ & 0.249 \\
$\kest_{\mathrm{global}}$ & 0.290 \\
$\hat{\mathbf{k}}_{p},~p=1$ & 0.423 \\ 
$\hat{\mathbf{k}}_p,~p=0.90$ & 0.458 \\ 
$\hatkmean$ & 0.501 \\ 
$\hat{\mathbf{k}}_p,~p=0.80$ & 0.505 \\ 
$\hat{\mathbf{k}}_p,~p=0.70$ & 0.560 \\ 
$\hatkvar$ & 0.990\\ \bottomrule
\end{tabular}
\label{tab:recap_estimates_metaSWE}
\end{table*}
Similarly as for $\JBH$ and $\JBHS$, the profiles are depicted Figure~\ref{fig:profiles_swe}.
\begin{figure}[!h]
  \centering
  \includegraphics[width=.8\textwidth]{Figures/profile_swe}
  \caption{Profiles for the calibration problem}
\label{fig:profiles_swe}
\end{figure}
We can see that the performances of $\kest_1$ and $\hatkmean$ are very similar, but $\kest_1$ has better performances when $\mathbf{u} > 0.6$.
\clearpage
\subsection{Assessing the quality of the forecast of the calibrated model}
The calibration of the model has been performed using by integrating the model on a time-period $[0, T]$, called assimilation window. 
We now want to compare the quality of the different forecasts, from different calibrated bottom friction.
Those forecasts result from the integration of the numerical model between the time $T$ and a time $T_{\mathrm{pred}}$.
